### 2014-11-19 doc/informe/paper.tex
actuazo tex

% Preview source code

%% LyX 2.1.1 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage[a4paper]{geometry}
\geometry{verbose,tmargin=1in,bmargin=1in,headheight=1in,headsep=1in}
\usepackage{float}
\usepackage{amsmath}
\usepackage{graphicx}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

\makeatother

\usepackage{babel}
\begin{document}

\title{A Pedestrian Simulationnnnnn}

\title{Future Virtual Particle Force Model}

\author{Castiglione Gonzalo, Agustin Marseillan, Daniel Parisi}
\maketitle
\begin{abstract}
We present a novel method to simulate virtual pedestrians based on
the social force model. In our model, each pedestrian has a point
in front of him called Future Virtual Particle (FVP) which represents
where the pedestrian is headed to and at what speed.
\end{abstract}
keywords:

pedestrian, collision avoidance, future virtual particle, force model

\pagebreak{}

\section{Introduction}

\subsection{Motivacion y antecedentes}

Navigation of biological, sinthetic or virtual agents is a relevant
problem in several fields such as pedestrian dynamics, moving robots
and animation of characters for videogames and motion pictures.

Modelling and simulating the displacement of agents througth arbitrarily
complex enviroments may be stated in an hierachical structure of mecahnisms
depending mianly on the distance from the agent. This level has been
named, from closer to further, as operational (walking, lowst level
physical-computaitnal model for displacement), tactical (wayfinding,
route chocie) and strategic (general activity planning) {[}Hoogonen
and Bovy 2004{]}. These levels are not independient, factors affecting
one level may impact in the following and vice versa, for example,
the route choice may vary due to congestion of agents producing from
previous route chice and walking behavior. Also, obstacles can impact
on the operational level or tactical level depending on the particular
geometry of the enviroment. The particular mechanism we want to address
is the avoidance of obstacles being fixed or moving (another agent)
which involves operational and tactical aspects of the navigation.

A general approach is to take an existing operational model and equip
it with a higher level model which allows better and smoother collision
avoidance behavior. Exisiting low level models can be taken from pedestrian
dynamics field and in general this models can be classified into rule
based and force based, discrete and continuous space description,
etc. {[}Schadschneider 2009{]}.

A famous example of continuos and force based model is the Social
Force Model {[}Helbing 1995, 2000{]}. In this model the dynamic for
virtual pedestrians is derived from the Newton equation\textquoteright s
considering the total force exceterd over each agent is the result
of three forces: Contact, Social and Driving Force. While the driving
force points towards the final objective of each pedestrian, the social
force is repulsive and acts as a kind of collision avoidance force.
However this social force term introduce several artefacts in some
configurations {[}see for example Lakoba et al. 2005, Parisi et al
2009{]}.

Cellular automata models make use of a spacial grid, which can be
occuied or empty, along with a set of rules determining the evolution
and conflict resolution of virtual pedestrians moving over the cells
of the grid. An emblematic celluar automata model is the one proposed
by Kirchner and Schadschneider {[}2002{]}.

It have been also proposed hibrid models such as the Contractile Particle
Model {[}Baglietto and Parisi, 2011{]} in which a continuos description
of the space is combined with a set of simple rules governig the dynamcis
of the system.

The basic operational model as the ones described above can be improved
if higher level mechanisms were added to manage more complex issues
as eficient avoidance. Some recent examples can be found in the literature.

Karamouzas et al. (2009) proposed a method for collision avoidance
modifying the social force model, basically, replacing the social
force term by a new \textquotedblleft evasive\textquotedblright{}
force which tends to avoid future collisions. The magnitud and direction
of this force is calculaed considering the predictions of these possible
collisions.

Kretz et al. 2011 have arrised the point that the key ingredient in
social force model is the driving force instead of interaction force,
so in this work the authors propose a method for dinamically adjust
the desired velocity following the gradient of a field given by a
time map, in other words, the desired velocity is choosen as the quickest
path to the objective taking into account the geometry and other agents
(collision, congestions, jams, etc.). Also mounted on the SFM, Moussaïd
et al. {[}2011{]} have presented a model using a \textquotedblleft cognitive
heuristics\textquotedblright{} to determine the norm and direction
of the desired velocity for each agent dinamically during the evolution
of the system.

In the same line, we also proposed that the navigation capacity of
virtual agents should be concentrated in the on-line decision of the
desired velocity. Similar to this approach, we postulate a new method
derived from the methods used in robotics/artificial inteligence\dots{}
The method deals with two aspects of the navigation mechanism. The
first one is geometrical problem of finding intermediate goals when
the final goal is not visible from the current position of the virtual
agent. The second aspect is the local avoidance of other agents, which
is performed via a reactive mechanism inspired in (robot navigation
/ artificial inteligence) aproaches ¿?¿?¿\dots ..?¿?¿ {[} Ref.???
??? {]}.

The method proposed could be mounted on different basic displacement
models like de SFM or the CPM, in the present work we have chosen
the first one.

\subsection{Social Force Model}

En el Social Force Model cada peatón se rige por una serie de reglas
presentads por Helbing en sus paper {[}Helbing 1995, 2000{]}. A partir
de estas reglas, proupso un modelo en donde cada peatón sufre una
fuerza de deseo que lo provoca a acercarse hacia su objetivo y una
fuerza de repulsión contra cada uno de los demas peatones. Junto a
estas dos, existe la fuerza de contacto producto del conectacto físico.
\\

Cada una de las fuerzas en el modelo se define usando las siguientes
fórmulas:\\

Fuerza de contacto (Granular):

\[
F_{G_{i}}=\sum_{j=1,j\ne i}^{N_{P}}[-\epsilon_{ij}k_{n}e_{ij}^{n}+v_{ij}^{t}\epsilon_{ij}k_{t}e_{ij}^{t}]\, g(\epsilon_{ij})
\]

En donde: $\epsilon_{ij}=r_{ij}-(R_{i}+R_{j})$

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.4]{sfm/ganular}
\par\end{centering}

\protect\caption{Fuerza de contact granular}

\end{figure}

\vspace{1cm}

Fuerza social:
\[
F_{S_{i}}=\sum_{j=1,j\ne i}^{N_{P}}Aexp(-\frac{\epsilon_{ij}}{B})\, e_{ij}^{n}
\]

\vspace{1cm}

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.6]{sfm/sf}
\par\end{centering}

\protect\caption{Fuerza social}
\end{figure}

Fuerza de deseo:

\[
F_{D_{i}}=m_{i}\frac{v_{di}e_{i}-v_{i}}{\tau}
\]

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.7]{sfm/drivingforce}
\par\end{centering}

\protect\caption{Driving force}
\end{figure}

\vspace{1cm}

Luego, se calcula $F_{i}=F_{G_{i}}+F_{S_{i}}+F_{D_{i}}$

En donde los parametros se encuentran definidos con lo siguientes
valores: $A=2000\,[N]$, $B=0.08\,[m]$, $k_{n}=1.2\,10^{5}\,[\frac{N}{m}]$,
$k_{t}=2.4\,10^{5}\,[\frac{kg}{m/s}]$ y $\tau=0.5\,[s]$.

\subsection{Future Vector Particle Model}

Debido a que el SFM presenta una fuerza ficticia sobre el peatón,
se generan artificios en la navegación ante la presencia de grandes
cantidades de peatones. \{...\}

Frente a esto se presenta un nuevo modelo

// no reproduce valores de Caudal Específico ni Diagrama Fundamental

// To do: explicar nuestros aportes al modelo, observaciones sobre
el SFM\\

\section{The Model}

\subsection{Hipotesis}

In the following the main effects that determine the motion of a pedestrian
will be introduced: (agregar que son las mismas hipotesis que presenta
Helbin en su paper)
\begin{enumerate}
\item The pedestrian wants to reach to his goal in the shortest possible
path.
\item The pedestrian's movement is influenced by other pedestrians. Depending
on the distance between the two of them and the predicted trayectory,
the pedestrian will feel the need to change his route to be able to
avoid the other pedestrian. It is because of this effect that pedestrians
will need to recalculate their route as new pedestrians get closer
to them.
\item Movement speed will be influenced by needs.
\end{enumerate}

\subsection{??}

A pedestrian is defined as follows:
\begin{itemize}
\item Circular shape

Represents the space that this pedestrian occupies. Circle's radio
is generated randomly to represent different types of pedestrians.
The range of values is distributed uniformly in $[0.25,0.29]$$[cm]$.

\item Long term objective

Represented by a static area. When it is touched by a pedestrian,
it is considered as accomplished. Multiple objectives can by defined
in a list, in this case, each of them must be reached in order.

\item Short term objective
\item Objetivo a corto plazo

Called FVP, it represents a point at a relative distance from the
pedestrian's center. It's a dynamic objective.

It is defined as a$1$ $[kg]$ mass. Not collisionable.

\item Desired speed

Represents the speed the pedestrian would walk if he was alone. Varies
randomly between$[1.2,1.4]$ $[m/s]$.

\item Reaction distance

Maximum distance between a pedestrian and his FVP, it represents the
distance at which a real pedestrian would react from an obstacle.

\end{itemize}
\vspace{1cm}

Fig 1: Pedestrian and his FVP.

\begin{figure}[H]
\centering{}\includegraphics[scale=0.5]{pedestrian-top}\protect\caption{Pedestrian top view}
\end{figure}

In the document, we will call pedestrians with a letter and FVP of
pedestrian x will be called x'.\\

\subsection{Implementation}

Each pedestrian has to reach the long term objective at some point,
to ensure this, the FVP always feels the need to be aligned with the
shortest path to the long term objective. On the other hand, there
are sometimes obstacles in the way, which will make this impossible,
in this cases, the route will have to change depending on the situation.

The pedestrian movement is calculated in four steps:
\begin{enumerate}
\item Calculate forces for each FVP.
\item Update positions for each FVP.
\item Calculate forces for each pedestrian.
\item Update positions for each pedestrian.\\

\end{enumerate}
Each step is defined as follows:
\begin{enumerate}
\item FVP force calculation.

In each iteration, the FVP from pedestrian $i$ feels a force generated
by the other pedestrians and FVPs. First of all, we filter de pedestrians
$j$ who are not in the range of sight of pedestrian $i$ as follows:

\[
k=i+\begin{pmatrix}0 & -1\\
1 & 0
\end{pmatrix}\overrightarrow{ii'}
\]
The following must be valid:

\[
(k_{x}-i_{x})*(j_{y}-i_{y})-(k_{y}-i_{y})*(j_{x}-i_{x})>0
\]

Once filtered, the force is calculated for every other FVP. The repulsion
force between i' and j' is defined as follows:

\[
F_{ext}(i)=\sum_{j}F_{i',j'}=\sum_{j}\alpha e^{-dist(i',j')/\beta}
\]

where $\alpha$ and $\beta$ are predefined constants. Afterwards,
we calculate the repulsion force between i' and j using the same formula
but different $\alpha$ and $\beta$. Afterwards, we calculate the
force between pedestrian i and his FVP i', this is a force that pushes
i' to be in its desired relative distance from the pedestrian. This
force is represented with a spring of length ``reaction distance''
and constant $K$. When we sum all these, we get the final $F_{ext}$.

Fig 2: FVP force calculation.

\begin{figure}[H]
\centering{}\includegraphics[scale=0.5]{pedestrian-top-forces}\protect\caption{Fuerzas que actúan sobre el Future $i'$}
\end{figure}

\begin{enumerate}
\item FVP position update.

In this stage, the position of the FVP is updated with Euler's formula
and the forces calculated in the previous step.

\end{enumerate}

To avoid high simetry situations, a low noise $P=10\%$ is added to
$F$. There are two ways to apply this noise:
\begin{itemize}
\item Radial noise:

\begin{itemize}
\item A value $p$ is taken randomly from a uniform distribution $[-P,\, P]$
and calculate: $FL_{i'}=F_{i'}*p$
\item Angular noise:

\begin{itemize}
\item A value $sgn=\{-1,\,1\}$ is taken randomly from a uniform distribution
and a value $p$ from $[-P,\, P]$. Then $FA_{i'}=rotation(F_{i'},\,\pi*sgn)*p$
is calculated.
\end{itemize}
\end{itemize}

At last, we find $F'_{i'}=F_{i'}+FL_{i'}+FA_{i'}$ and apply euler's
movement equations.

\end{itemize}
\begin{enumerate}
\item Pedestrian forces calculation.

The pedestrian always wants to move in the direction his FVP is pointing
and its magnitude is defined as Fd or desire force:

\[
F_{deseo}=\frac{dist(i',i)}{dist_{react}}\frac{v_{desire}-v_{curr}}{\tau}\frac{i'-i}{|i'-i|}
\]

Where $\tau=0.5$

\item Pedestrian position update.

The new position of the pedestrian will be calculated using Euler's
formula and the desire force calculated in the previous step.

\end{enumerate}
\end{enumerate}
TODO: Los valores que se usaron son $\alpha=800$ y $\beta=[0.65,\,085]$
con distribución uniforme.

\section{Calibracion}

\subsection{Metricas a usar}

Los resultados obtenidos fueron comparados utilizando el modelo ``Social
Force Model'' utilizando los valores propuestos por Dirk Helbing
{[}2{]}. Los escenarios de prueba fueron cruce y hallway.

Los resultados fueron cuantificados utilizando las siguientes métricas:
\begin{enumerate}
\item Cantidad total de colisiones

TODO: (como se calcula)

\item Duración promedio de las coliciones

TODO: (como se calcula)

\item Velocidad promedio de viaje

TODO: (como se calcula)

\item Tiempo promedio de viaje

TODO: (como se calcula)

\item Promedio de la distancia recorrida

TODO: (como se calcula)

\item Ángulo giro promedio

TODO: (como se calcula)

\end{enumerate}

\section{Results}

\vspace{1cm}
En la tabla a continuación se presentan las mediciones de las métricas
utilizando un total de $10$ corridas:

\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
Valores & 1 & 2 & 3 & 4 & 5 & 6\tabularnewline
\hline 
\hline 
$\alpha=800,\,\beta=[0.65,\,0.85]$ & (1.800, 0.748)  & (34.600, 8.333)  & (1.024, 0.016)  & (1.910, 0.022)  & (2.392, 0.005)  & (111.062, 19.122)\tabularnewline
\hline 
$\alpha=2000,\,\beta=0.08$ & (5.333, 2.625) & (12.333, 8.340)  & (1.052, 0.003)  & (1.876, 0.003)  & (2.391, 0.002)  & (106.185, 13.778)\tabularnewline
\hline 
\end{tabular}

\textbf{// Poner Graficos indicando distancias y esquemas del future
y la particula. }

\section{Conclusiones}

\textbf{// agregar al final futuras opciones que se abren con este
trabajo}

\pagebreak{}

\section{Referencias}
\begin{itemize}
\item {[}1{]} ... Karamouzas ...
\item {[}2{]} ...Helbing ...\end{itemize}

\end{document}