### 2015-01-24 doc/informe/paper.tex
actualizo carpeta informe con todo lo que esta subido en shareslatex

\documentclass[english]{article}
\usepackage[latin9]{inputenc}
\usepackage[a4paper]{geometry}
\geometry{verbose,tmargin=1in,bmargin=1in,headheight=1in,headsep=1in}
\usepackage{float}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{graphicx}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
\numberwithin{equation}{section}
\numberwithin{figure}{section}

\makeatother

\usepackage{babel}
\begin{document}

\title{Future Virtual Particle Method for Pedestrian Navigation}

\author{Castiglione Gonzalo, Agustin Marseillan, Daniel Parisi}
\maketitle
\begin{abstract}
We present a novel method to simulate virtual pedestrians based on
the social force model. In our model, each pedestrian has a point
in front of him called Future Virtual Particle (FVP) which represents
where the pedestrian is headed to and at what speed.
\end{abstract}
keywords:

pedestrian, collision avoidance, future virtual particle, force model

\pagebreak{}

\section{Introduction}

\subsection{Motivation and previous work}

Navigation of biological, sinthetic or virtual agents is a relevant
problem in several fields such as pedestrian dynamics, moving robots
and animation of characters for videogames and motion pictures.

Modelling and simulating the displacement of agents through arbitrarily
complex environments may be stated in an hierarchical structure of mechanisms
depending mainly on the distance from the agent. This level has been
named, from closer to further, as operational (walking, lowest level
physical-computational model for displacement), tactical (way-finding,
route choice) and strategic (general activity planning) {[}Hoogonen
and Bovy 2004{]}. These levels are not independent, factors affecting
one level may impact in the following and vice-versa, for example,
the route choice may vary due to congestion of agents produced from
previous route choice and walking behavior. Also, obstacles can impact
on the operational level or tactical level depending on the particular
geometry of the environment. The particular mechanism we want to address
is the avoidance of obstacles being fixed or moving (another agent)
which involves operational and tactical aspects of the navigation.

A general approach is to take an existing operational model and equip
it with a higher level model which allows better and smoother collision
avoidance behavior. Existing low level models can be taken from pedestrian
dynamics field and in general this models can be classified into rule
based and force based, discrete and continuous space description,
etc. {[}Schadschneider 2009{]}.

A famous example of continuous and force based model is the Social
Force Model {[}Helbing 1995, 2000{]}. In this model the dynamic for
virtual pedestrians is derived from the Newton equation\textquoteright s
considering the total force exerted over each agent is the result
of three forces: Contact, Social and Driving Force. While the driving
force points towards the final objective of each pedestrian, the social
force is repulsive and acts as a kind of collision avoidance force.
However this social force term introduces several artifices in some
configurations {[}see for example Lakoba et al. 2005, Parisi et al
2009{]}.

Cellular automaton models make use of a spatial grid, which can be
occupied or empty, along with a set of rules determining the evolution
and conflict resolution of virtual pedestrians moving over the cells
of the grid. An emblematic cellular automaton model is the one proposed
by Kirchner and Schadschneider {[}2002{]}.

Hybrid models have also been proposed such as the Contractile Particle
Model {[}Baglietto and Parisi, 2011{]} in which a continuous description
of the space is combined with a set of simple rules governing the dynamics
of the system.

The basic operational model -as the ones described above- can be improved
if higher level mechanisms were added to manage more complex issues
as efficient avoidance. Some recent examples can be found in the literature.

Karamouzas et al. (2009) proposed a method for collision avoidance
modifying the social force model, basically, replacing the social
force term by a new \textquotedblleft evasive\textquotedblright{}
force which tends to avoid future collisions. The magnitude and direction
of this force is calculated considering the predictions of these possible
collisions.

Kretz et al. 2011 have arrised the point that the key ingredient in
social force model is the driving force instead of interaction force,
so in this work the authors propose a method for dynamically adjusting
the desired velocity following the gradient of a field given by a
time map, in other words, the desired velocity is chosen as the quickest
path to the objective taking into account the geometry and other agents
(collision, congestion, jams, etc.). Also mounted on the SFM, Moussaïd
et al. {[}2011{]} presented a model using \textquotedblleft cognitive
heuristics\textquotedblright{} to determine the norm and direction
of the desired velocity for each agent dynamically during the evolution
of the system.

In the same line, we also proposed that the navigation capacity of
virtual agents should be concentrated in the on-line decision of the
desired velocity. Both direction and magnitude calculation of the desired velocity
are the key differences with the SFM.

The method proposed could be mounted on different basic displacement
models like the SFM or the CPM, in the present work we have chosen
the first one.

\subsection{Social Force Model}

//Hay que citar la de 1995 pero la que estamos describiendo es la del 2000
{[}Helbing 1995, 2000{]} presents a Model in which each pedestrian
is subject to ``social forces''. This forces are a measure for the
internal motivations of the individual to perform certain actions.
In its simplest form, the Social Force Model consist on three main
terms. Each of them represents a force that a pedestrian will
suffer in order to reach his objective.

The first term describes the acceleration that a pedestrian keeps
towards his desired velocity of motion. This term is known as the
``Driving Force'' and it's calculated as follows:

\begin{equation}
\label{eq:driving-force}
F_{D_{i}}=m_{i}\frac{v_{di}e_{i}-v_{i}}{\tau}
\end{equation}

In this latest formula, $v_{di}$ represents the desire speed for
the pedestrian $i$, $v_{i}$ is the current velocity and $e_{i}$
represents the desired direction of motion of the pedestrian $i$
and is calculated as follows:

\begin{equation}
\label{eq:desired-direction}
e_{i}(t)=\frac{r_{i}^{k}-r_{i}(t)}{||r_{i}^{k}-r_{i}(t)||}
\end{equation}

where $r_{i}(t)$ denotes the actual position of the pedestrian
$i$ at the time $t$ and $r_{i}^{k}$ is the closest point from the goal
(represented as an area) to the pedestrian. The unit vector $e_{i}$ will 
be pointing towards the closest point. Figure \ref{fig:driving-force} shows an 
example of a pedestrian currently moving in direction of $v$ but correcting its 
trajectory towards $x$.

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.4]{pics/sfm/drivingforce}
\par\end{centering}

\protect\caption{\label{fig:driving-force}Driving force}
\end{figure}

\vspace{1cm}

The second term represents the fact that a pedestrian keeps a certain
distance to other pedestrians and borders. This term is known as the
``Social Force'' and it's calculated as follows:

\[
F_{S_{i}}=\sum_{j=1,j\ne i}^{N_{P}}Aexp(-\frac{\epsilon_{ij}}{B})\, e_{ij}^{n}
\]

In this formula the terms $A$ and $B$ represent constants that
were determined by simulations and $\epsilon_{ij}$ represents the
distance from pedestrian $i$ towards pedestrian $j$. Figure $1.2$
shows this vector graphically.

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.4]{pics/sfm/sf}
\par\end{centering}

\protect\caption{Social Force}
\end{figure}

\vspace{1cm}

The third term represents the granular force. This force represents
the physical force that a pedestrian feels when colliding with
another object (pedestrian or wall). This granular force term is
defined as follows:

\[
F_{G_{i}}=\sum_{j=1,j\ne i}^{N_{P}}[-\epsilon_{ij}k_{n}e_{ij}^{n}+v_{ij}^{t}\epsilon_{ij}k_{t}e_{ij}^{t}]\, g(\epsilon_{ij})
\]

Where: $\epsilon_{ij}=r_{ij}-(R_{i}+R_{j})$

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.4]{pics/sfm/ganular}
\par\end{centering}

\protect\caption{Example of two pedestrians in collision}

\end{figure}

\vspace{1cm}

Afterwards $F_{i}=F_{G_{i}}+F_{S_{i}}+F_{D_{i}}$is calculated for
each of the pedestrians on each simulation step and applied to the
pedestrian until it reaches his goal.

Fixed parameter values: $A=2000\,[N]$, $B=0.08\,[m]$, $k_{n}=1.2\,10^{5}\,[\frac{N}{m}]$,
$k_{t}=2.4\,10^{5}\,[\frac{kg}{m/s}]$ y $\tau=0.5\,[s]$.

While this model doesn't present a very real behaviour for pedestrians,
it worked as a starting point for numerous projects.

\subsection{Future Virtual Particle Model}

Given that the SFM adds a fictional force on pedestrians, navigation doesn't resemble reality for high pedestrian densities.
Also, SFM didn't have the same values as well known metrics for real-case scenarios such as the flow of pedestrians going out a door and the fundamental diagram.

Because of this, we present a new model in which the social force from equation \ref{eq:driving-force} affecting directly on each pedestrian is eliminated and replaced with a desire force facing a short term objective. This force is created using equation \ref{eq:desired-direction} replacing $r_{i}^{k}$ with the vector facing the closest point on the short term objective area.

\section{The Model}

\subsection{Hipotesis}

The main effects that govern the motion of a pedestrian are the same
as Helbin's:
\begin{enumerate}
\item The pedestrian wants to reach his goal in the shortest possible path.
\item The pedestrian's movement is influenced by other pedestrians. Depending
on the distance between the two of them and the predicted trajectory,
the pedestrian will feel the need to change his route to be able to
avoid the other pedestrians. It is because of this effect that pedestrians
will need to recalculate their route as new pedestrians get closer
to them.
\item Movement speed will be influenced by needs.
\end{enumerate}

\subsection{Geometrical definition}

A pedestrian is defined as follows:
\begin{itemize}
\item Circular shape

Represents the space that this pedestrian occupies. Circle's radio
is generated randomly to represent different types of pedestrians.
The range of values is distributed uniformly in $[0.25,0.29]$$[cm]$.

\item Long term objective

Represented by a static area. When it is touched by a pedestrian,
it is considered as accomplished. Multiple objectives can by defined
in a list, in this case, each of them must be reached in order.

\item Short term objective

Called future virtual particle (FVP), it represents a point at a
relative distance from the pedestrian's center. It's a dynamic objective.

It is defined as a $1\,[kg]$ mass. Not collisionable.

\item Desired speed

Represents the speed the pedestrian would walk if he was alone. Varies
with a uniform distribution between $[1.2,1.4]\,[m/s]$.

\item Reaction distance ($RD$)

Maximum distance between a pedestrian and his FVP, it represents the
distance at which a real pedestrian would react from an obstacle.

\end{itemize}
\vspace{1cm}

Figure \ref{fig:Pedestrian-top-view} shows a pedestrian geometrical
definition:

\begin{figure}[H]
\centering{}\includegraphics[scale=0.5]{pics/pedestrian-top}\protect\caption{\label{fig:Pedestrian-top-view}Pedestrian top view}
\end{figure}
\vspace{1cm}
Pedestrians will be named after a letter (i.e: $i$ or $j$) and the same letter with a apostrophe will represent it's FVP (i.e: $i'$ or $j'$)\\

Figure \ref{fig:Pedestrian-vectors} shows all the vectorial definition
that will be used:

\begin{figure}[H]
\centering{}\includegraphics[scale=0.5]{pics/geometry}\protect\caption{\label{fig:Pedestrian-vectors}Vectorial definitions}
\end{figure}

\vspace{1cm}

\subsection{Dynamics of the FVP Model}

In this section all the forces acting on the FVPM will be defined

\subsubsection{Forces}
\begin{itemize}
\item Dynamic of the FVP

Each pedestrian has to reach the long term objective at some point,
to ensure this, the FVP needs to be aligned with the
shortest path to the long term objective $\mathbf{o}$. On the other hand, there
are sometimes obstacles in the way, which will make this impossible,
in this cases, the route will have to change depending on the situation.

To model this situations, two types of forces act over the FVP:

\begin{itemize}
\item Internal force
\\
\\
This force will guide the pedestrian on the shortest path $\mathbf{ox_i}$ and minimize the time needed to reach the long term goal. This force is modeled using two springs:
\\
\iffalse
This forces will try to keep the FVP aligned with $\mathbf{o}$ and at the same time, 
keep the FVP at $RD$ from the pedestrian.

This (first) spring will regulate the velocity
that will be calculated on a posterior step at which the pedestrian
$i$ will advance. 
...
which pushes the FVP towards its desired length.
At the same time, a force pushing the FVP towards his original
alingment, defined

\fi

The first spring starts on $x_{i'}$ and ends on the point at distance $RD$ on vector $\mathbf{x_{i}o}$ with a spring constant $K_{1}$.
This spring represents how much the pedestrian wants to reach his target.
A larger spring constant will force a more straight path, but possible
with more collisions on its way.
In order to avoid an oscillatory movement (i.e: with a single pedestrian), a 
damping $\gamma$ was added to this spring.

The second spring starts on $x_i$, ends on $x_i'$ and has a steady 
distance of $RD$ with a spring constant $K_s(\theta)$ 
where $\theta$ is the angle between $\mathbf{u}$ and $\mathbf{r}$.
With this spring the FVP will tend to always be at $RD$ distance 
from the pedestrian.
Because a pedestrian always tries to reach his goal in the shortest
possible path (hypothesis 1), if it has to take a big detour of his
ideal path, it will try to reduce his velocity drastically in order
to avoid making a long travel. To recreate this, the
spring constant has to be dependant of the deviation angle. 
\[
K_s(\theta) = \frac{K_{2}}{\theta}
\]

Given pedestrian $i$, the equation for his internal force is:

\begin{equation}
\iffalse
F_s(r,\theta)=K_s(\theta)\frac{\vec{ii'}}{\mid\vec{ii'}\mid}(r_{max}-r_{ii'})
\fi
F_{internal}(i)=K_{1}*(|r|-RD)+K_{2}(\theta_{i})+K_{2}(\theta_{i})*(???)
\end{equation}

\begin{figure}[H]
\centering{}\includegraphics[scale=0.5]{pics/pedestrian-internal-forces}\protect\caption{Pedestrian internal forces}
\end{figure}

\item Avoidance forces
\\
\\
This force will produce avoidance movements when obstactes are detected on the path.
The magnitude of this force is calculated using two factors:
\\
The first factor is calculated for each of the pedestrian $j$ ($j\neq i$)
who is in the range of sight of pedestrian $i$. This restriction
is verified using the following condition:

\[
\mathbf{r_{ii'}}\bullet\mathbf{r_{ij'}}<0
\]

This filter represents the fact that a pedestrian (in most cases)
does not make decisions based on obstacles behind him. Hence,
they are ignored. The formula for the external forces that the
FVP $i$ feels is defined as follows

\[
F_{ext}(i)=\sum_{j}F_{i',j'}+F_{i'j}=\sum_{j}(\alpha_{ff}e^{-dist(i',j')/\beta_{ff}})+\sum_{j}(\alpha_{fp}e^{-dist(i',j')/\beta_{fp}})
\]

where $\alpha_{x}$ and $\beta_{x}$ are predefined constants.

The first term of the sum acts as a repulsion force between $i$ and $j$
FVPs, resulting in the avoidance of a future collision.
The second term uses this same principle but calculates the repulsion
force between $j's$ FVP and $i$.
Repulsion against walls is calculated in the same way, by using the
closest point between the FVP and the wall and special values for
$\alpha_{fw}$ and $\beta_{fw}$ for the calculation.

\iffalse
The first term of the sum represents
the fact that $i$ does not want to be in the same location as $j$
in the near future. Note that $j'$ represents the predicted location
of $j$ inferred by another pedestrian (in this case, $i$) and because
of this, it moves it's near future away of that predicted location.
The second term, represents the same fact but taking in mind the current
location of the pedestrian $j$. The same law applies to each of the
walls on the simulation, the difference is that the closest point
on the wall and and the future has to be calculated before applying
the formula. Also, the constants $\alpha_{fw}$ and $\beta_{fw}$
have different values.
\fi

Figure \ref{fig:External-forces-affecting} shows the external forces
that a pedestrian $i$ suffers because of another pedestrian ($j$)
and also the direction in which $i$ desires to move.

\begin{figure}[H]
\centering{}\includegraphics[scale=0.5]{pics/pedestrian-top-forces}\protect\caption{\label{fig:External-forces-affecting}External forces affecting future
$i'$}
\end{figure}

\end{itemize}

After this, the final $F_{i'}$ is computed by adding all the terms.
$F_{i'}=F_{ext}(i)+F_{i\, s1}+F_{i\, s2}$.\\

To avoid high symetry situations, a low noise $P=10\%$ is added to
$F$. There are two ways to apply this noise:
\begin{itemize}
\item Radial noise:

\begin{itemize}
\item A value $p$ is taken randomly from a uniform distribution $[-P,\, P]$
and calculate: $FL_{i'}=F_{i'}*p$
\item Angular noise:

\begin{itemize}
\item A value $sgn=\{-1,\,1\}$ is taken randomly from a uniform distribution
and a value $p$ from $[-P,\, P]$. Then $FA_{i'}=rotation(F_{i'},\,\pi*sgn)*p$
is calculated.
\end{itemize}
\end{itemize}

At last, we find $F'_{i'}=F_{i'}+FL_{i'}+FA_{i'}$ and apply movement
equations.

\end{itemize}
\item Dynamic of the pedestrian

The pedestrian always wants to move in the direction his FVP is pointing
and its magnitude is defined as $F_{d}$ or desire force:

\[
F_{desire_{i}}=m_{i}\frac{\frac{|\vec{ii'}|}{|\vec{r}_{max}|}\vec{e_{i}}-\vec{v}_{i}}{\tau}
\]

Where $\tau=0.5$

\textit{*It is important to note that the only deviation from the SFM in this
equation is the $k\vec{e}_i$ term. As described above, this model focuses on improving
the SFM through dynamically adjusting the desired velocity.}

\end{itemize}

\subsubsection{Algorithm}

The pedestrian movement is calculated in four steps:
\begin{enumerate}
\item Calculate forces for each FVP.
\item Update positions for each FVP.
\item Calculate forces for each pedestrian.
\item Update positions for each pedestrian.
\end{enumerate}

\section{Calibration}

\subsection{Metrics}

The results where compared to the SFM using the values proposed by
Dirk Helbing {[}2{]}. The test scenarios where crossing and hallway
for they present the main two types of symmetry (90 degrees and 180
degrees).

The metrics used where:
\begin{enumerate}
\item Number of collisions

When a pedestrian's body is touching another one, a counter is incremented,
it won't be incremented again unless it separates and touch again.
This counter represents the amount of collisions.

\item Total duration of collisions:

When a pedestrian's body is touching another one, a counter is incremented
by the time step of the simulation. This counter will represent the
total time pedestrians were colliding.

\item Average walking speed:

//TODO

\item Average travel time:

When a pedestrian appears in the map, this time is recorded, once
he reaches his goal, the travel time is calculated with this value.
The average of all this times represents the average travel time.

\item Average travel distance:

Each step, the distance a pedestrian has traveled from the source
is saved. When this pedestrian reaches his goal, that distance is
considered as finished and it is saved elsewhere. Afterwards, an average
of all finished distances is calculated.

\item Average turn angle:

Each step, the angle of the previous velocity of a pedestrian and
the current one is saved and added to a total. When this pedestrians
reaches his goal, this total is considered as finished and it is saved
elsewhere. Afterwards, an average of all finished turn angles is calculated.

\end{enumerate}

\subsection{Values}

To calibrate the model, runs varying parameters were made. A wide
spectrum of values was covered, testing every combination of every
possible one. After seeing clear preferences towards certain values,
the values were refined within that scope. After numerous iterations
of this process, the values that best suit these metrics are$\alpha=800$
y $\beta=[0.65,\,085]$ uniformly distributed.

//TODO todos los demas parametros

\section{Results}

\vspace{1cm}

\begin{table}[H]
\centering{}{\scriptsize{}}%
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
{\scriptsize{}Values} & {\scriptsize{}1} & {\scriptsize{}2} & {\scriptsize{}3} & {\scriptsize{}4} & {\scriptsize{}5} & {\scriptsize{}6}\tabularnewline
\hline 
\hline 
{\scriptsize{}$\alpha=1000,\,\beta=[0.4,\,0.5]$} & {\scriptsize{}(1.800, 0.748) } & {\scriptsize{}(34.600, 8.333) } & {\scriptsize{}(1.024, 0.016) } & {\scriptsize{}(1.910, 0.022) } & {\scriptsize{}(2.392, 0.005) } & {\scriptsize{}(111.062, 19.122)}\tabularnewline
\hline 
{\scriptsize{}$\alpha=2000,\,\beta=0.08$} & {\scriptsize{}(5.333, 2.625)} & {\scriptsize{}(12.333, 8.340) } & {\scriptsize{}(1.052, 0.003) } & {\scriptsize{}(1.876, 0.003) } & {\scriptsize{}(2.391, 0.002) } & {\scriptsize{}(106.185, 13.778)}\tabularnewline
\hline 
\end{tabular}\protect\caption{Metrics comparing SFM vs FVPM. Average of $10$ runs.}
\end{table}

\textbf{// Poner Graficos indicando distancias y esquemas del future
y la particula. }

\section{Conclusions}

\textbf{// agregar al final futuras opciones que se abren con este
trabajo}

\pagebreak{}

\section{References}
\begin{itemize}
\item {[}1{]} ... Karamouzas ...
\item {[}2{]} ...Helbing ...\end{itemize}

\end{document}